---
title: "Important codes"
author: "Humphrey"
date: "2022-07-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Learn `group_by` from the dplyr package

group_by() allows us to group on a categorical column.

I count() the presence of automobile classes by manufacturer.

Then ungroup() to remove any leftover groups. 

And voila, I now have counts by mfger + class.

```{r}
library(dplyr)
library(ggplot2)

mpg %>% 
  group_by(manufacturer) %>% 
  count(class, name="n") %>% 
  ungroup()
```

### But you can simplify with just `count()`

```{r}
mpg %>% 
  count(manufacturer, class, name = "n")
```

## Learn `pivot_wider()` from tidyr

pivot_wider() is a great function (replaces spread() which I also enjoy using)

And we can now make pivot tables that shift our values from the column class into new columns. 

Super valuable for creating those pivot table reports.

```{r}
library(tidyr)
mpg_pivot_table1 <- mpg %>% 
  group_by(manufacturer) %>% 
  count(class, name = "n") %>% 
  ungroup() %>% 
  pivot_wider(
    names_from = class,
    values_from = n,
    values_fill = 0
  )
mpg_pivot_table1
```

## Learn pivot_longer() from tidyr

Pivot longer reverses the operation by taking a wide data frame and pivoting it into a long "tidy" format.

The "long format" is super important for making visualizations with ggplot2 & doing iteration with purrr::map().

```{r}
mpg_long_summary_table <- mpg_pivot_table1 %>% 
  pivot_longer(
    cols = compact:subcompact,
    names_to = "class",
    values_to = "value"
  )
mpg_long_summary_table
```
## Learn how to combine group_by() + summarise() + across()

I love this quick 3-punch combo. ❤️

group_by() for setting up groups

summarise() for applying summary functions to each group

across() for applying multiple functions to one or more columns


