---
title: "Important codes"
author: "Humphrey"
date: "2022-07-07"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

## Learn `group_by` from the dplyr package

-   `group_by()` allows us to group on a categorical column.

-   I `count()` the presence of automobile classes by manufacturer.

-   Then `ungroup()` to remove any leftover groups.

-   And now have counts by manufacturer + class.

```{r}
library(dplyr)
library(ggplot2)

mpg %>% 
  group_by(manufacturer) %>% 
  count(class, name="n") %>% 
  ungroup()
```

### But you can simplify with just `count()`

```{r}
mpg %>% 
  count(manufacturer, class, name = "n")
```

## Learn `pivot_wider()` from `tidyr`

-   `pivot_wider()` is a great function (replaces `spread()` which I also enjoy using)

-   And we can now make pivot tables that shift our values from the column class into new columns.

-   Super valuable for creating those pivot table reports.

```{r}
library(tidyr)
mpg_pivot_table1 <- mpg %>% 
  group_by(manufacturer) %>% 
  count(class, name = "n") %>% 
  ungroup() %>% 
  pivot_wider(
    names_from = class,
    values_from = n,
    values_fill = 0
  )
mpg_pivot_table1
```

## Learn `pivot_longer()` from `tidyr`

-   Pivot longer reverses the operation by taking a wide data frame and pivoting it into a long "tidy" format.

-   The "long format" is super important for making visualizations with ggplot2 & doing iteration with `purrr::map()`.

```{r}
mpg_long_summary_table <- mpg_pivot_table1 %>% 
  pivot_longer(
    cols = compact:subcompact,
    names_to = "class",
    values_to = "value"
  )
mpg_long_summary_table
```

## Learn how to combine `group_by()` + `summarise()` + `across()`

-   `group_by()` for setting up groups

-   `summarise()` for applying summary functions to each group

-   `across()` for applying multiple functions to one or more columns

```{r}
mpg %>%
  group_by(class) %>% 
  summarise(
    across(cty, .fns = list(mean=mean, stdev=sd)), .groups = "drop"
  )
```

## Learn `relocate()` from dplyr

-   `relocate()` allows us to have complete control on how we move columns.

-   I can still use `select()` for this, but when I absolutely need fine control, I switch over to `relocate()`

```{r}
mpg %>% 
  relocate(where(is.character), .after = last_col())
```

## Learn `group_split()` from dplyr

-   I prefer group_split() for splitting data frames that are grouped into a list containing sub-data frames as elements.

    *You'll see why in the next tip.*

**Note** *(Pro-Tip)*: Make sure to convert your grouping column into a factor first, which preserves the order

```{r}
df_split <- mpg %>% 
  mutate(manufacturer=as.factor(manufacturer)) %>% 
  group_by(manufacturer) %>% 
  group_split()
df_split[[1]]
```

### We can now combine `group_split()` with `map()`

-   This is a cool example where I'm splitting the data frame by manufacturer then applying a linear regression model to each data frame.

-   I get a linear reg model for each car manufacturer.

-   print out the linear reg model for the first 2 car manufacturers

```{r}
library(purrr)
#library(kableExtra)
linear_reg_models <- mpg %>% 
  mutate(manufacturer=as.factor(manufacturer)) %>% 
  group_by(manufacturer) %>% 
  group_split() %>% 
  
  map(.f=function(df){
  lm(hwy~cty, data = df)
}
  )
linear_reg_models[c(1:2)]
```
